---
title: "yli130 Assignment_3"
author: "Yanxi Li"
date: "10/16/2020"
output: html_document
---

### NaiveBayes Assignment

***

### Data Preparation
Use only 5 categorical variables for the predictor variable. 
Transform DAY_WEEK, DEP_TIME and Flight.Status into factors.
Round DEP_TIME and label Flight.Status into 0 = on time, 1 = delayed.
```{r}
data <- read.csv("FlightDelays.csv") # import data from desktop
head(data)
library(dplyr)
colnames(data)
my_data <- select(data, CARRIER, DEP_TIME, DEST, ORIGIN, DAY_WEEK, Flight.Status)    # choose 5 categorical variables
head(my_data)
str(my_data)     
my_data$DAY_WEEK <- factor(my_data$DAY_WEEK)   # transform day_week into factors
my_data$DEP_TIME <- round((my_data$DEP_TIME)/100)   # round departure time 
my_data$DEP_TIME <- factor(my_data$DEP_TIME)  # transform dep_time into factors
my_data$Flight.Status <- factor(my_data$Flight.Status,   
                                levels = c("ontime", "delayed"),
                                labels = c(0, 1))   # transform flight.status into factor and label it
str(my_data)
```

***

### Divide into 60% train and 40% validation 
We do not need to normalize the categories because they are not numerical.
```{r}
library(caret)
set.seed(123)
Index_train <- createDataPartition(my_data$CARRIER, p=0.6, list = FALSE)  # Train data is 60%
Train_data <- my_data[Index_train,]   
Validation_data <- my_data[-Index_train,]   # Validation data is rest (40%)
```

***

### Use Naivebayes model to predict
Use library e1071 or naivebayes to predict whether the flight is delayed or not.
```{r}
library(e1071)
# build a naivebayes classfier
model <- naiveBayes(Flight.Status ~., data = Train_data)
Predict_validation_data <- predict(model, Validation_data)
head(Predict_validation_data)
Predict_validation_data_prob <- predict(model, Validation_data, type = "raw")
head(Predict_validation_data_prob)
```

***

### Counts table & Proportion table
```{r}
# counts table 
mytable <- table(my_data$Flight.Status, my_data$ORIGIN)  # row is flight status and column is destiantion airport
mytable
# proportion table   
prop.table(mytable, 2)   # 2 means calculate the column percentages
```

***

### Confusion matrix & ROC Curve
```{r}
library(gmodels)
CrossTable(x = Validation_data$Flight.Status, y = Predict_validation_data, prop.chisq = FALSE)  # Accuracy = 0.827
library(pROC)
roc(Validation_data$Flight.Status, Predict_validation_data_prob[,1])    # AUC(area under the curve) value
plot.roc(Validation_data$Flight.Status, Predict_validation_data_prob[,1])
```


***

### END




















